\section{Component systems}
The most common example of such systems is a set of books. In this case one puts on the rows the words in the whole vocabulary and the books' titles on the columns. The entry that corresponds to row $i$ and column $j$ is the number of times the word $i$ appears in the book $j$. The same happens if one considers Wikipedia's pages. Other examples are: Lego$\textsuperscript{\tiny\textregistered}$ sets where components are the Lego$\textsuperscript{\tiny\textregistered}$ bricks and realizations the Lego$\textsuperscript{\tiny\textregistered}$ packages, and protein domains; all these were described and well studied in~\cite{mazzolini2018heaps, Mazzolini2018zipf}.

Given a matrix with $N$ components on the rows and $R$ realizations on the columns and relative abundances $n_{ij}$ as the entries, it is interesting to study some quantities that are universal and general characteristics of component systems.

First of all, the \textbf{occurrence} of a component is defined as 
\begin{equation}\label{eq:occurrence}
O_i=\frac{\sum_{j=1}^{R}(1-\delta_{n_{ij},0})}{R}.
\end{equation}
It is the fraction of realizations in which the component's abundance is not null. A component that is present in all the realizations has got $O_i=1$, the ensemble of the components with $O_i=1$ is known as the \textbf{core}. Components with high ($\simeq 1$) occurrence are present in mostly all realizations of the datasets. In linguistics articles, such as \textit{the}, are present everywhere, so they have high occurrence. Components with low occurrence ($\simeq 0$) are present only in a few realizations and are the most specific ones~\cite{altmann2016statistical}.

The sum across all columns, or the number of times component $i$ appear in the dataset, is called \textbf{abundance} of the component and is defined as
\begin{equation}\label{eq:abundance}
a_i=\sum_{j=1}^{R}n_{ij};
\end{equation}
dividing this by the global abundance, or the total number of components in the dataset,
\begin{equation}
  a=\sum_{i=1}^{N}a_i
\end{equation}
naturally brings to the \textbf{frequency of a component} in the whole corpus
\begin{equation}\label{eq:fi}
f_i=\frac{a_i}{\sum_{c=1}^{N}a_{c}}.
\end{equation}
The abundance of a component divided by the sum of all the abundances in a realization gives the \textbf{frequency} of the component in that specific realization
\begin{equation}
f_{ij}=\frac{n_{ij}}{\sum_{c=1}^{N}n_{cj}}.
\end{equation}

The sum of all abundances in a realization,
\begin{equation}\label{eq:size}
M_j=\sum_{c=1}^{N}n_{cj}
\end{equation}
represents the \textbf{size} of the realization. In gene expression this is the size of the transcript.

It is expected that frequencies distribute according to the so-called Zipf's law
\begin{equation}\label{eq:zipf}
f_i\propto r_i^{-\alpha}
\end{equation}
where $r$ is the rank: the position of a component when sorting, in descending order, all components by their frequencies in the whole dataset.