\chapter*{Abstract}
The interest in studying complex systems is increasingly spreading.
Complex systems can be found anywhere and many common behaviours are observable among them, systems with different origins and purposes may share  some statistical laws.

An example can be the Zipf's law, well-known in linguistics and texts analysis. It can be easily observed in the distribution of gene expression in different samples of cancer tissues.

In recent years, datasets with a large amount of cancer samples' data are available, the most complete is The Cancer Genome Atlas (TCGA). From this dataset, it is easy to obtain, for example, gene expression data from RNA-sequencing experiments together with a lot of metadata about the samples themselves. Another dataset containing healthy tissues (GTEx) will be analysed for comparison and benchmark.

If one studies the number of samples in which a gene is expressed above a certain threshold, the so-called occurrence, it is easily verified that there are different kinds of genes. Some are present in the majority of samples, some others are present only in a subset of the whole dataset. The same behaviour can be found analysing words in a corpus of texts; some words, such as \emph{the}, are present everywhere, other specific words are present only in texts regarding a certain
subject. This suggests that there are similarities between a system of words and documents and a system of genes and samples.

Given a corpus of documents, these can be classified by their specific subject. Similarly, samples can be classified, for example, by the tissue they come from or by the type of the disease they are referred to.

The similarities between gene expression and linguistics data suggest the possibility to use topic modelling to classify data and separate samples and genes in different clusters. Topic modelling is a set of clustering algorithms in networks' theory. Given a set of words and documents, these algorithms describe documents as a mixture of topics. Topics are nothing but communities of words each one with a given probability.

Purpose of this work is to build a bipartite network of genes and samples and use topic modelling to find communities. The goal is to separate samples depending on the site the tumour was and the disease type of the sample. Moreover, it is possible to separate genes depending on their specific functions. Once a community structure of genes emerges, it is possible to run a hypergeometric test on the whole set to verify if it reveals some type of enrichment and to inspect common properties among genes.

The specific algorithm used in this work is particularly unique because it needs no priors and makes no assumptions on the data. Moreover, it can be set to accept overlapping clusters so it is possible to find genes belonging to different topics and it can be hierarchic. All these facts empower a lot of new possibilities to investigate a network.

A hierarchic approach makes it possible to classify data at different layers. An ideal goal would be to separate healthy and diseased samples at the first layer, then separate them by tissue, then by the tumour type and so on.
